The yaq Project: \\ Standardized Software Enabling Flexible Instrumentation

### 2022-11-11 paper.tex
minor changes

The Wright Group at UW-Madison needs to orchestrate a large variety of hardware in multidimensional scans for their complex spectroscopy experiments \cite{MukamelShaul2000a, WrightJohnCurtis2011a}.
This need for exquisite hardware control has resulted in several prior attempts at ``home-built'' orchestration software \cite{CarlsonRogerJohn1988a, MeyerKentAlbert2004b, KainSchuyler2017a, ThompsonBlaiseJonathan2018a, SundenKyle2022}.
Now, using \yaq{}, the Wright Group has been able to move to Bluesky rather than inventing their own sophisticated control software ``from scratch''.
The Wright Group uses simulated hardware to enable client development away from the active laboratory computers.
This allows Wright Group researchers to create polished client interfaces without interrupting  ongoing experiments.
Clients developed by the Wright Group have proved flexible enough to be used on four laser systems, each with different complements of hardware.
Moving forward, the Wright Group will spend less energy developing control software and more energy developing creative spectroscopy experiments.

@@

In a \yaq{} context, new hardware can be incorporated into an instrument through the addition of a new daemon.
The \yaq{} architecture simplifies hardware interface development in several ways.
First, because daemons are separate and portable programs, the development effort can be spread across the community of \yaq{} users.
Often, researchers can download an existing daemon rather than writing a new one.
Second, \yaq{} daemon development can be performed separated from the particulars of any individual client.
Often, this means that initial hardware enablement work can be done on a researchers personal machine before the new hardware peripheral is installed in the instrument.
Third, when developing trait-compliant protocols, it becomes easy to design and fully test your hardware interface.
Traits are unambiguous and well-described, making an obvious target for development.
Tooling exists to verify full trait compliance, for example you can use \texttt{yaqc-qtpy} to provide a graphical program to interact with your hardware immediately.
Finally, as discussed in Section II, \yaq{} gives you options to design using remote hardware or unusual interfaces when necessary.

@@

In \yaq{}, each component of an instrument can be developed and distributed separately.
For example, two different instruments might happen to use the same temperature sensor.
Because the temperature sensor daemon is its own independent program, both instruments can benefit from the same daemon.
The growing ``ecosystem'' of \yaq{} daemons make future instruments easier and easier to develop.
Growing this ecosystem is a collaborative effort where many \yaq{} users create portable daemons that they need and share them with the community where they will be used and improved.
\yaq{} components can be incorporated seamlessly into existing software projects, including languages other than Python.
For example, an existing LabVIEW project may add additional hardware support via \yaq{}.
In this way, a more modular instrument can be built gradually without needing to overhaul all of the software simultaneously.

>>>>

The Wright Group at UW-Madison needs to orchestrate a large variety of hardware in multidimensional scans for their complex spectroscopy experiments \cite{MukamelShaul2000a, WrightJohnCurtis2011a}.
This need for exquisite hardware control has resulted in several prior attempts at ``home-built'' orchestration software \cite{CarlsonRogerJohn1988a, MeyerKentAlbert2004b, KainSchuyler2017a, ThompsonBlaiseJonathan2018a, SundenKyle2022}.
Now, using \yaq{}, the Wright Group has been able to move to Bluesky rather than inventing their own sophisticated control software ``from scratch''.
The Wright Group uses simulated hardware to enable client development away from the active laboratory computers.
This allows Wright Group researchers to create polished client interfaces without interrupting  ongoing experiments.
Clients developed by the Wright Group have proved flexible enough to be used on four laser systems, each with different complements of hardware.
The Wright Group expects to spend less energy developing control software and more energy developing creative spectroscopy experiments in future work.

@@

In a \yaq{} context, new hardware can be incorporated into an instrument through the addition of a new daemon.
The \yaq{} architecture simplifies hardware interface development in several ways.
First, because daemons are separate and portable programs, the development effort can be spread across the community of \yaq{} users.
Often, researchers can download an existing daemon rather than writing a new one.
Second, \yaq{} daemon development can be performed separated from the particulars of any individual client.
Often, this means that initial hardware enablement work can be done on a researchers personal machine before the new hardware peripheral is installed in the instrument.
Third, when developing trait-compliant protocols, it becomes easy to design and fully test your hardware interface.
Traits are unambiguous and well-described, making an obvious target for development.
Tooling exists to verify full trait compliance, for example you can use \texttt{yaqc-qtpy} to provide a graphical program to interact with your hardware immediately.
Finally, as discussed in Section II, \yaq{} provides options to design using remote hardware or unusual interfaces when necessary.

@@

In \yaq{}, each component of an instrument can be developed and distributed separately.
For example, two different instruments might happen to use the same temperature sensor.
Because the temperature sensor daemon is its own independent program, both instruments can benefit from the same daemon.
The growing ``ecosystem'' of \yaq{} daemons make future instruments easier and easier to develop.
Growing this ecosystem is a collaborative effort where many \yaq{} users create portable daemons that they need and share them with the community where they will be used and improved.
\yaq{} components can be incorporated seamlessly into existing software projects, including languages other than Python.
For example, an existing LabVIEW project may provide additional hardware support via \yaq{}.
In this way, a more modular instrument can be built gradually without needing to overhaul all of the software simultaneously.
